Make any stupid plot you want

### 2021-01-18 draft/sections/intro.tex
q
raided my second exam for data + rv intro to lead into math data

>>>>

set up: dubois

theorists:
bertin, munzner, mackinlay

talk about:
matplotlib arch paper, excel/matlab arch, 
vtk \& ggplot (compare/contrast, we're blending these things)

\subsection{Terminology}
\begin{figure}[h!]
%%remake properly at some point
\includegraphics[width=\textwidth]{figures/intro/flowchart.png} 
\caption{To go from a dataset to a visualization, the data is subset based on a set of constraints (the invariant). The resulting subset becomes the components that are visualized, but the choice of visualization is dependent on the type and structure of the component variables.}
\label{fig:flowchart}
\end{figure}

Given a dataset, we need to decide what subset of the data to visualize. Bertin describes the set of constraints used to subset the data as the \textit{invariant}. Formally, the \textit{invariant} is the set of shared characteristics of the data being visualized. When these constraints are applied to the dataset, the resulting subset is what will become the \textit{components} of the visualization \cite{bertin_semiology_2011}. In figure~\ref{fig:iris_scatter}, the \textit{invariant} common to all the data being visualized is "sepal length", "petal length", and "species" and the \textit{components} are the measurements of these variables. As shown in figure~\ref{fig:flowchart}, the final step in creating a visualization is choosing how to encode the
components using retinal (visual) variables.

\subsubsection{Retinal (Visual) Variables}
\begin{figure}[h!]
\includegraphics[width=1\textwidth]{figures/intro/retinal_variables.png}
\caption{Retinal variables are a codification of how position, size, shape, color and texture are used to illustrate variations in the \textit{components} of a visualization. This tabular form of Bertin's retinal variables is from Understanding Graphics \cite{_information_????} who reproduced it from \textit{Making Maps: A Visual Guide to Map Design for GIS} \cite{krygier_making_2005}}
\label{fig:retinal_variables}
\end{figure}

Figure~\ref{fig:retinal_variables} illustrates common guidelines for encoding \textit{components}, derived from what Bertin terms a retinal variable and most other visualization theorists call visual variables \cite{bertin_semiology_2011,krygier_making_2005,chambers_graphical_1983,wilkinson_grammar_2005,munzner_visualization_2014}. The columns of figure~\ref{fig:retinal_variables} correspond to the type of observation: discrete points, continuous events (e.g. a timeseries), two dimensional continuous events (e.g. a vector map). The rows of figure~\ref{fig:retinal_variables} describe ways to visualize variations in the \textit{components}; generally, quantitative components are represented by retinal variables that change quantitatively and categorical components are represented by retinal variables that vary qualitatively. In figure~\ref{fig:iris_scatter}, the hue of the marker is used to encode differentiation in species and the position of the marker is used to show variation in petal length and sepal length. The retinal variables suggest that any single visualization is limited to encoding at most about 8 or 9 components. Retinal variables provide guidelines for encoding \textit{components}, but the choice of graph is based on the type and structure of the data.

\subsubsection{Data Type and Structure}

\begin{figure}[h!]
 \includegraphics[width=\textwidth]{figures/intro/munzner_datatypes}
\caption{Keys are unique lookup values used to find individual observations in the dataset. Keys are positional references, and can be coordinates on a map or unique values such as a primary key in a database or a (time, latitude, longitude) index in a data cube. Image modified from a diagram from Munzner's website \cite{_visualization_????}}
\label{fig:munzner_datatypes}
\end{figure}

As shown in figure~\ref{fig:flowchart}, there are multiple ways to translate data into pictures. A map is always an option, except when the observations do not have associated coordinates in a physical plane. Tamara Munzner provides a way to distinguish between these datasets using $\{key, value\}$ designations \cite{munzner_visualization_2014}. Munzner defines \textit{values} as measurements of interest in the dataset, analogous to dependent variables in statistics. She defines \textit{keys} as indexes that can be used to look up values, analogous to independent variables in statistics and dimensions in computer science. Figure~\ref{fig:munzner_datatypes} illustrates how these keys are used to look up variables in a dataset: 
\begin{itemize}
	\item map: keys are the coordinates of the points
	\item table: row index, database primary key
	\item data cube: row, column, etc. (.e.g. $i,j,k$) index
\end{itemize}

Expanding on Munzner's key and value semantics, in many datasets the keys are discrete variables like time or geophysical locations sampled from a continuous curve, surface, or field. While these observations are discrete samples from the continuous space, often the continuous (functional) characteristic\cite{ramsay_functional_2006,muller_functional_2006} of the observational space is also of interest. Besides quantitative discrete, quantitative continuous, or categorical measurement type considerations, the choice of visualization is also influenced by the measurement being on an interval, ratio, nominal, or categorical scale.

### 2021-01-18 draft/sections/math.tex
raided my second exam for data + rv intro to lead into math data

In this section, we describe the fiber bundle spaces of data(\ref{sec:data}), graphics(\ref{sec:graphic}), and intermediate visual characteristics (\ref{sec:artist}). We then discuss the equivariant maps between data and visual characteristics (~\ref{sec:artist_nu}) and visual characteristics and graphics (~\ref{sec:artist_q}) that make up the artist.

\subsection{Data Space}
\label{sec:space_data}
% introduce what this is gonna happen in this section - we will talk about, fiber and connectivity 
% motivate w/ data/ example - have weather station. %tidydata 
% we will introduce the fiber bundle and its components base and total space to explain the fiber

@@

%%% might get moved into basespace discussion:

We model data as a fiber bundle ($E$, $K$, $\pi$, $F$)
\begin{equation}
    \label{eq:fiber_bundle}
    \begin{tikzcd}
        F \arrow[r, hook] & E \arrow[r, "\pi"] & K
    \end{tikzcd}
\end{equation}

\subsubsection{Variable Field in F }
\label{sec:data}
 %%%first sentence should be a paragraph
%%% butler's paper, no notion of distance on K, can only map F to the visual fibers, so need an X to map to Xpos
% variable, measurement, pull in tidy here, 
% % motivate by example, temperature \ time 
%% F_0 = {temp_0, temp_n}, 
%% then adding in a second set of variables time: 
%% F_1 = {time_0, time_1}
% break out into F first as set of values, then F=F_{0}\times F_{1} as component

@@

%%close out the fiber bundle section with the cartesian products, which gives us heterogenous data , univariate to multivariate. We can decompose the fiber bundle to components bundles \pi:E_1\rightarrow K, .... \pi: E_n\rightarrow K
\subsubsection{Connectivity: Base Space} 
%% flows out of the fiber section into, now gluing observations together into a thing
%% can then put functions on edges/planes/etc so we have true continouity

\label{sec:base_data}
\begin{figure}[h!]
    \includegraphics[width=\textwidth]{figures/math/k_different_types.png}
    \label{fig:base_space_types}
    \caption{The topological base space $K$ encodes the connectivity of the data space, for example if the data is independent points or a map or on a sphere}
\end{figure}

@@

\subsubsection{Data: Section}
%%% visual variable encoding in a consistent/atomic way in a complex strcutrue 
\label{sec:section_data}
The sections of the fiber bundles are the instances of the data. They are the functions that map from a point $k \in K$ to a set of values in $F$. For example, if we have a database, the section is the table that yields a row given an index. The section $\tau$ is the mapping from base space to total space $\tau: K\rightarrow E$ 
\begin{equation}
    \label{eq:section}
    \begin{tikzcd}
        F \arrow[r, hook] & E \arrow[d, "\pi"']           \\
        & K \arrow[u, "\tau"', bend right]
    \end{tikzcd}
\end{equation}

@@

\subsubsection{Visual Idioms: Equivalance class of artists}
As formulated above, every artist function $A$ has fixed $\nu$ and generates a distinct graphic $\rho$. It is unfeasible to implement $A$ for every single graphic; instead we implement the equivalence class of artists $\{A \in A^\prime: A_{1} \equiv A_{2}\}$ which is $Q:\Gamma(V) \rightarrow \Gamma(H)$. 
\end{document}

>>>>

We first describe the fiber bundle spaces of data(\ref{sec:data}), graphics(\ref{sec:graphic}), and intermediate visual characteristics (\ref{sec:artist}). We then discuss the equivariant maps between data and visual characteristics (\ref{sec:artist_nu}) and visual characteristics and graphics (\ref{sec:artist_q}) that make up the artist.

\subsection{Data Space}
\label{sec:data}
Generally when we discuss data, we are interested in two aspects: the variables in the dataset

% introduce what this is gonna happen in this section - we will talk about, fiber and connectivity 
% motivate w/ data/ example - have weather station. %tidydata 
% we will introduce the fiber bundle and its components base and total space to explain the fiber

@@

\subsubsection{Variable Field in F }
\label{sec:data_fiber}
 %%%first sentence should be a paragraph
%%% butler's paper, no notion of distance on K, can only map F to the visual fibers, so need an X to map to Xpos
% variable, measurement, pull in tidy here, 
% % motivate by example, temperature \ time 
%% F_0 = {temp_0, temp_n}, 
%% then adding in a second set of variables time: 
%% F_1 = {time_0, time_1}
% break out into F first as set of values, then F=F_{0}\times F_{1} as component

@@

%%close out the fiber bundle section with the cartesian products, which gives us heterogenous data , univariate to multivariate. We can decompose the fiber bundle to components bundles \pi:E_1\rightarrow K, .... \pi: E_n\rightarrow K
\subsubsection{Connectivity: Base Space} 
\label{sec:data_base}
%% flows out of the fiber section into, now gluing observations together into a thing
%% can then put functions on edges/planes/etc so we have true continouity
We model data as a fiber bundle ($E$, $K$, $\pi$, $F$)
\begin{equation}
    \label{eq:fiber_bundle}
    \begin{tikzcd}
        F \arrow[r, hook] & E \arrow[r, "\pi"] & K
    \end{tikzcd}
\end{equation}

\begin{figure}[h!]
    \includegraphics[width=\textwidth]{figures/math/k_different_types.png}
    \label{fig:base_space_types}
    \caption{The topological base space $K$ encodes the connectivity of the data space, for example if the data is independent points or a map or on a sphere}
\end{figure}

@@

\subsubsection{Data: Section}
%%% visual variable encoding in a consistent/atomic way in a complex structure 
\label{sec:section_data}
The sections of the fiber bundles are the instances of the data. They are the functions that map from a point $k \in K$ to a set of values in $F$. For example, if we have a database, the section is the table that yields a row given an index. The section $\tau$ is the mapping from base space to total space $\tau: K\rightarrow E$ 
\begin{equation}
    \label{eq:section}
    \begin{tikzcd}
        F \arrow[r, hook] & E \arrow[d, "\pi"']           \\
        & K \arrow[u, "\tau"', bend right]
    \end{tikzcd}
\end{equation}

@@

\subsubsection{Visual Idioms: Equivalance class of artists}
As formulated above, every artist function $A$ has fixed $\nu$ and generates a distinct graphic $\rho$. It is unfeasible to implement $A$ for every single graphic; instead we implement the equivalence class of artists $\{A \in A^\prime: A_{1} \equiv A_{2}\}$ which is $Q:\Gamma(V) \rightarrow \Gamma(H)$.

%% Well maybe. I think this is where required components comes in. There is some minimal V where if you didn't have those variables you wouldn't call it a scatter plot .... something like xpos,ypos with K just a descrete set.
%% so V probably can be any V that contains that sub-bundle
%% so if bubble plot is not the same as scatter plot, and something where the colors change based on data is not scatter plot then it is the equivalent class of something with that minimal bundle.
%% But then if you want to include ANYTHING that also maps with X-Y ... then you hit "direct limits" where you limit over all possivle V_big containing V_minimal
%% You still get a map on some A_minimal -> A_big
%% If you know how to visualize with more variables .. you can always fix them.
\end{document}
